---
title: "ESLR 2024 Multi-agent Reinforcement Learning with a Static Graph"
output: html_document
date: "`r Sys.Date()`"
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
  
  )
```

```{css, echo=FALSE}
.output{
  background-color:white;
  border: 1px solid black;
  font-weight: bold;
}
```

```{r class.source="output",eval = FALSE, include = FALSE}
alpha_base <- convert_prob_to_base(alpha)
```

```{r, echo=FALSE}
library(ggplot2)
library(igraph)
library(tidyverse)
library(shiny)
library(rstudioapi)

```

# Social Learning

```{r,  out.width='50%', fig.align='center', fig.cap='classic social learning',  echo=FALSE}
fileP = getSourceEditorContext()$path
getP <- gsub("ESLR202402.Rmd", "individual.png", fileP)
knitr::include_graphics(getP)
```

<br>

<br>

```{r, out.width='50%', fig.align='center', fig.cap='collective social learning',  echo=FALSE}
fileP = getSourceEditorContext()$path
getP <- gsub("ESLR202402.Rmd", "collective.jpg", fileP)
knitr::include_graphics(getP)
```

<br>

## Question

#### If multiple agents learn simultaneously, how can we quantify this dynamic process?

<br>

# Outline

-   Reinforcement Learning

-   Network

-   Simulating Multi-agent Reinforcement Learning

# Reinforcement Learning

### Individual RL

<br> Reinforcement learning concerns a family of problems in which an agent evolves while analyzing consequences of its actions, based on a simple scalar signal (the reinforcement) derived from the environment. The purpose of reinforcement learning is for the agent to learn an optimal (or near-optimal) policy that maximizes the reward function or other user-provided reinforcement signal that accumulates from immediate rewards. <br>

<br>

```{r, out.width='70%', fig.align='center', fig.cap='The agent–environment interaction in a Markov decision process',  echo=FALSE}
fileP = getSourceEditorContext()$path
getP <- gsub("ESLR202402.Rmd", "RL.png", fileP)
knitr::include_graphics(getP)
```

<br>

1 - at time step *t*, the agent is in state $S_{t}$,

2 - agent chooses one of the possible actions in this state, $A_{t}$,

3 - agent applies the action, what provokes,:

-   the passage to a new state, $S_{t+1}$,

-   the receipt of the reinforcement, $R_{t}$;

4 - *t* \<- *t + 1*

5 - go to 2 or stop if the new state is a terminal one.

<br>

**Two most important distinguishing features of reinforcement learning:**

-   Trial-and-error search

-   Delayed reward

<br>

<br>
**Social reinforcement in complex contagion**
```{r, out.width='40%', fig.align='center',   echo=FALSE}
fileP = getSourceEditorContext()$path
getP <- gsub("ESLR202402.Rmd", "compexcontagion.png", fileP)
knitr::include_graphics(getP)
```

<br>

### k-armed bandit problem

<br>

```{r, out.width='20%', fig.align='center',   echo=FALSE}
fileP = getSourceEditorContext()$path
getP <- gsub("ESLR202402.Rmd", "bee.png", fileP)
knitr::include_graphics(getP)
```

::::: {style="display: flex;"}
<div>

```{r, out.width='25%', fig.cap = "Patch01", fig.align='center',   echo=FALSE}
fileP = getSourceEditorContext()$path
getP <- gsub("ESLR202402.Rmd", "patch01.jpg", fileP)
knitr::include_graphics(getP)
```

```{r, echo=FALSE}

ui <- fluidPage(
  
  
   sidebarLayout(

    # Sidebar with a slider input
    sidebarPanel(
  
    sliderInput("mean_adjust", label = "Mean adjustment:",
              min = 1, max = 8, value = 2, step = 0.5),
  
  sliderInput("sd_adjust", label = "sd adjustment:",
              min = 0.01, max = 10, value = 4, step = 0.01),
  actionButton("update", "Forage!")),
   mainPanel(
  plotOutput("plot"),
  textOutput("reward")
   )
))



server <- function(input, output) {
  output$plot <- renderPlot({
  randomData <- rnorm(10000, mean = input$mean_adjust, sd = input$sd_adjust)
  hist(randomData, freq = TRUE,  breaks = 50,  xlim = c(0,10), ylim = c(0,1000),xlab = "Reward", ylab = "Frequency",  main = "Patch01 reward")
})

output$reward <-renderText({ 
    randomData <- rnorm(10000, mean = input$mean_adjust, sd = input$sd_adjust)
    sample(randomData,1)
  }) |>
    bindEvent(input$update)

}

shinyApp(ui = ui, server = server, options = list(height = 800, width = 500))

```

</div>

<div>

```{r, out.width='25%', fig.cap = "Patch02", fig.align='center',   echo=FALSE}
fileP = getSourceEditorContext()$path
getP <- gsub("ESLR202402.Rmd", "patch02.jpg", fileP)
knitr::include_graphics(getP)
```

```{r, echo=FALSE}

ui <- fluidPage(
  
  
   sidebarLayout(

    # Sidebar with a slider input
    sidebarPanel(
  
    sliderInput("mean_adjust", label = "Mean adjustment:",
              min = 1, max = 8, value = 6, step = 0.5),
  
  sliderInput("sd_adjust", label = "sd adjustment:",
              min = 0.01, max = 1, value = 0.2, step = 0.01),
  actionButton("update", "Forage!")),
   mainPanel(
  plotOutput("plot"),
  textOutput("reward")
   )
))



server <- function(input, output) {
  output$plot <- renderPlot({
  randomData <- rnorm(10000, mean = input$mean_adjust, sd = input$sd_adjust)
  hist(randomData, freq = TRUE,  breaks = 50,  xlim = c(-10,10), ylim = c(0,1000),xlab = "Reward", ylab = "Frequency",  main = "Patch02 reward")
})

output$reward <-renderText({ 
    randomData <- rnorm(10000, mean = input$mean_adjust, sd = input$sd_adjust)
    sample(randomData,1)
  }) |>
    bindEvent(input$update)

}

shinyApp(ui = ui, server = server, options = list(height = 800, width = 500))

```

</div>
:::::

<br> 


## Task setup

```{r, out.width='70%', fig.cap = "collective foraging setup", fig.align='center',   echo=FALSE}
fileP = getSourceEditorContext()$path
getP <- gsub("ESLR202402.Rmd", "task.png", fileP)
knitr::include_graphics(getP)
```









<br>

### Model-free reinforcement learning: Q-learning

In reinforcement learning, learning is essentially the evaluation of actions (not from correct instructions). Q-learning is based on the Temporal Differences and used to determine an optimal policy. It doesn't require the knowledge of probability transitions from a state to another and is model-free.

<br>

**basic Q learning Model**

$$
Q_{s,a,t}\leftarrow Q_{s,a,t} + \alpha(r_{a,t}+\gamma\max a Q_{a,t}-Q_{a,t})
$$

<br>

$Q_{a,t}$ current value

$\alpha$ a step-size parameter influences learning rate $0 \leq \alpha \leq 1$

$\gamma$ discount factor $0\leq\gamma<1$

$\gamma$ discounts the value of future rewards compared to immediate rewards. A higher discount factor means that future rewards are more valuable, encouraging long-term beneficial actions over short-term gains. $gamma$ = 0, agent is myopic and only cares about the immediate reward. $gamma$ approaching 1, agent looks more ahead.

<br>

**Value function: Rescorla-Wagner rule** 

<br>

$$
Q_{a,t+1 }\leftarrow Q_{a,t } + \alpha\pi_{a,t}
$$

<br>

$$\pi_{a,t}\leftarrow r_{a,t}-Q_{a,t}$$

$\pi_{a,t}$ prediction error 

<br>

<br>

**From value to action: Action selection function**

Simplest action selection: *greedy* action selection

<br>

$$
A_{t}\doteq \arg\max_{a} Q_{a,t}
$$ $\epsilon-greedy$ method

$$
   A_{t}\doteq 
   \left\{ \begin{array}{rcl}
        \arg\max_{a} Q_{a,t} & 1-\epsilon,
         \\ random & \epsilon.
         \end{array}\right.
$$

Asocial:*softmax* function

```{r, out.width='60%', fig.align='center', fig.cap='Sofmax function',  echo=FALSE}
fileP = getSourceEditorContext()$path
getP <- gsub("ESLR202402.Rmd", "softmax.png", fileP)
knitr::include_graphics(getP)
```

### Multi-agent Reinforcement Learning

Social: frequency-dependent strategy 

<br>

$$
P_{a,t+1 } = (1-\sigma)\frac{exp(\beta Q_{a,t})}{\sum exp(\beta Q_{k,t})}+\sigma\frac{N_{a,t}^\theta}{\sum N_{k,t}^\theta}
$$ 

<br> 

$\beta$ inverse temperature: stochasticity of action(exploration-exploitation tendency)

$\sigma$ copying weight

$\theta$ conformity bias(strength of conformity)

**Extended model:** **between states State-Value Function**

<br>

```{r, out.width='60%', fig.align='center', fig.cap='Policy evaluation',  echo=FALSE}
fileP = getSourceEditorContext()$path
getP <- gsub("ESLR202402.Rmd", "maze.png", fileP)
knitr::include_graphics(getP)
```

<br> 
Bellman Equation 

describes the relationship between a current state's value and future states' values. <br>



<br> 

$$
V_{s} = \max_{a}(R_{s,a}+\gamma\sum_{s^{'}} P(s^{'}|s,a)V_{s^{'}})
$$

<br> Q Value is updated: <br>

$$
Q_{s,a,t+1 }\leftarrow r_{s,a,t} + \gamma\sum s^{'}P(s^{'}|s,a) max a^{'}Q_{s^{'},a^{'},t}
$$ 
Extended reading for details: Reinforcement Learning: An Introduction, 2nd edition by Richard S. Sutton and Andrew G.

<br>

<br>

# Network

<br>

<br>

```{r, out.width='50%', fig.cap = "Patial information in collective behavior", fig.align='center',   echo=FALSE}

fileP = getSourceEditorContext()$path
getP <- gsub("ESLR202402.Rmd", "partial.jpg", fileP)
knitr::include_graphics(getP)
```

<br>

<br>

## Using network to represent diverse social interactions

**Full graph**

```{r, include = TRUE}

g <- make_full_graph(40)
plot(g, vertex.size=6, vertex.label=NA)
```

**Tree graph**

```{r, include = TRUE}
treeg <- make_tree(50, children = 8, mode = "undirected")

plot(treeg, vertex.size=6, vertex.label=NA) 
```

**Star graph**

```{r, include = TRUE}
star <- make_star(40)
plot(star, vertex.size=6, vertex.label=NA) 
```

**Erdos-Renyi random graph** $G(n,p)$

```{r, eval = FALSE}
#$G(n,p)$ graph
gnp <- sample_gnp(n=100, p=.005)
sphere <- layout_(gnp, on_sphere())
plot(gnp, vertex.size=6, vertex.label=NA, layout = sphere)
```

```{r, eval = FALSE, include = FALSE}
# gnp.rewired <- rewire(gnp, each_edge(prob=0.1))
# plot(gnp.rewired, vertex.size=6, vertex.label=NA, layout = sphere)
#
```

<br> 
<br>

::::: {style="display: flex;"}
<div>

```{r, echo=FALSE}

ui <- fluidPage(


  actionButton("update", "Give me a Randim Graph!"),
#
  plotOutput("plot", height = 450, width = 450)
  # plotOutput("rewirep", height = 450, width = 450)

)

# ui <- fluidPage(
#   
#   
#    sidebarLayout(
# 
#     # Sidebar with a slider input
#     sidebarPanel(
#   
#   actionButton("update", "Rewire!")),
#    mainPanel(
#   plotOutput("plot", height = 450, width = 450),
#   plotOutput("rewirep", height = 450, width = 450)
#    )
# ))
# 


server <- function(input, output) {
library(igraph)
  

    output$plot <- renderPlot({

   gnp <- sample_gnp(n=100, p=.005)
   sphere <- layout_(gnp, on_sphere())
   plot(gnp, vertex.size=6, vertex.label=NA, layout = layout.sphere)

  })|>
     bindEvent(input$update)

 
#    result <- eventReactive(input$update, {
#   
#    gnp <- sample_gnp(n=100, p=.005) 
#    gnp.rewired <- rewire(gnp, each_edge(prob=0.1))
#    })
#  
# output$rewirep <-renderPlot({
#   
#    rewirep <- plot(result(), vertex.size=6, vertex.label=NA, layout = sphere)
#    
#   }) |>
#     bindEvent(input$update)

# output$plot <-renderPlot({
#   
#    gnp <- sample_gnp(n=100, p=.005) 
#    sphere <- layout_(gnp, on_sphere())
#    plot <-plot(gnp, vertex.size=6, vertex.label=NA, layout = sphere) 
#   })
# 
# 

}
shinyApp(ui = ui, server = server, options = list(height = 500, width = 450))
#shinyApp(ui = ui, server = server, options = list(height = 800, width = 500))

```

</div>

<div>

```{r, echo=FALSE}

ui <- fluidPage(


  actionButton("update", "Rewire!"),
#
# plotOutput("plot", height = 450, width = 450),
  plotOutput("rewirep", height = 450, width = 450)

)

# ui <- fluidPage(
#   
#   
#    sidebarLayout(
# 
#     # Sidebar with a slider input
#     sidebarPanel(
#   
#   actionButton("update", "Rewire!")),
#    mainPanel(
#   plotOutput("plot", height = 450, width = 450),
#   plotOutput("rewirep", height = 450, width = 450)
#    )
# ))

server <- function(input, output) {
library(igraph)
  
  
  #   output$plot <- renderPlot({
  #      
  #  gnp <- sample_gnp(n=100, p=.005)
  #  sphere <- layout_(gnp, on_sphere())
  #  plot(gnp, vertex.size=6, vertex.label=NA, layout = sphere) 
  #   
  # })
  # 
 
   result <- eventReactive(input$update, {
  
   gnp <- sample_gnp(n=100, p=.005) 
   gnp.rewired <- rewire(gnp, each_edge(prob=0.1))
   })
 
output$rewirep <-renderPlot({
  
   rewirep <- plot(result(), vertex.size=6, vertex.label=NA, layout = layout.sphere)
   
  }) |>
    bindEvent(input$update)

# output$plot <-renderPlot({
#   
#    gnp <- sample_gnp(n=100, p=.005) 
#    sphere <- layout_(gnp, on_sphere())
#    plot <-plot(gnp, vertex.size=6, vertex.label=NA, layout = sphere) 
#   })
# 
# 

}
shinyApp(ui = ui, server = server, options = list(height = 500, width = 450))
#shinyApp(ui = ui, server = server, options = list(height = 800, width = 500))

```

</div>
:::::

<br> 
<br>

## Simulating Multi-agent Reinforcement Learning

### Network setup

Let's use the simple Erdos-Renyi random graph $G(n,p)$ as a demonstration.

**Erdos-Renyi random graph** $G(n,p)$

Get the adjacency matrix for later social frequency calculation

```{r, include = TRUE}
# generate a G(n,p) graph
gnp <- sample_gnp(n = 10, p = .5)

# adjacency matrix
gnp_adjMatrix <- get.adjacency(
  gnp,
  type = c("both"),
)
gnp_adjMatrix 
plot(gnp, vertex.size=6, vertex.label=NA, layout = layout.sphere)

```

```{r, include = FALSE}
# generate a G(n,p) graph

gnp <- sample_gnp(n = 10, p = .25)
plot(gnp, vertex.size=6, vertex.label=NA, layout = layout.sphere)
```

### Global setup

```{r, include = TRUE}
# Session Setup
set.seed(6) 
group_size = 10 
repetition = 10 # number of repetition for a specific combination of parameters
horizon = 100 # total steps of one session


# Bandit Setup
num_options = 2 # the number of arms
mean_list = c(4, 5) 
sd_list = c(1, 1)


# RL parameter space
Q_initial = 4.5 # prior belief
alpha_list <- c(0.3, 0.5, 0.7)
beta_list <- c(2, 4, 6)
sigma_list <- seq(0, 1, 0.2)
theta_list <- seq(1, 7, 2)


```

<br> 
**Container ready for data recording**

```{r, include = TRUE}


social_learning_model_data <- 
  data.frame(
    
    beta = rep(beta_list, times = 1, each = length(alpha_list)*length(sigma_list)*length(theta_list)*repetition*horizon)
    , alpha = rep(alpha_list, times = length(beta_list), each = length(sigma_list)*length(theta_list)*repetition*horizon)
    , sigma = rep(sigma_list, times = length(beta_list)*length(alpha_list), each = length(theta_list)*repetition*horizon)               
    , theta = rep(theta_list, times = length(beta_list)*length(alpha_list)*length(sigma_list), each = repetition*horizon)               
    , r = rep(1:repetition, times = length(beta_list)*length(alpha_list)*length(sigma_list)*length(theta_list), each = horizon)
    , t = rep(1:horizon, times = length(beta_list)*length(alpha_list)*length(sigma_list)*length(theta_list), each = 1)
    , bestChoiceProb = rep(NA, length(beta_list)*length(alpha_list)*length(sigma_list)*length(theta_list)*repetition*horizon)
    , averageReward = rep(NA, length(beta_list)*length(alpha_list)*length(sigma_list)*length(theta_list)*repetition*horizon)
  )


```

<br> 
**Individual differences**

Noise add to learning parameters.Later parameter recovery test.

Fix $\alpha$ to 0.3

<br>

```{r, eval = FALSE}

 # Setting individual learning parameters
      alpha_base <- convert_prob_to_base(alpha)
      this_alpha <- (alpha_base + rnorm(group_size, 0, 0.05)) %>% convert_base_to_prob
      this_beta <- beta + rnorm(group_size, 0, 0.05)
      
```

```{r, echo = FALSE }
library(tidyverse)
alpha = 0.3
beta = 6
group_size = 10


## Converting function
convert_prob_to_base = function (alpha) {
    alpha[which(alpha < 1e-5)] <- 1e-5
    alpha[which(alpha > 1 - 1e-5)] <- 1 - 1e-5
    return (log(-alpha/(alpha-1)))
}

convert_base_to_prob = function (alphaRaw) {
    alphaRaw[which(alphaRaw > 12)] <- 12
    return( 1 / (1 + exp(-alphaRaw)))
}


 # Setting individual learning parameters
      alpha_base <- convert_prob_to_base(alpha)
      this_alpha <- (alpha_base + rnorm(group_size, 0, 0.05)) %>% convert_base_to_prob
      this_alpha
    
```

<br>

Fix $\beta$ to 6

<br>

```{r, eval = FALSE}

      this_beta <- beta + rnorm(group_size, 0, 0.05)
      
```

```{r, echo = FALSE }
library(tidyverse)
alpha = 0.3
beta = 6
group_size = 10


## Converting function
convert_prob_to_base = function (alpha) {
    alpha[which(alpha < 1e-5)] <- 1e-5
    alpha[which(alpha > 1 - 1e-5)] <- 1 - 1e-5
    return (log(-alpha/(alpha-1)))
}

convert_base_to_prob = function (alphaRaw) {
    alphaRaw[which(alphaRaw > 12)] <- 12
    return( 1 / (1 + exp(-alphaRaw)))
}


 # Setting individual learning parameters
      
      this_beta <- beta + rnorm(group_size, 0, 0.05)
     
      this_beta
```

<br>

### Container setup

**Value**

Setup initial Q value

```{r, eval = FALSE}

      Q <- array(dim = c(num_options, horizon, group_size))
      Q[,1,] <- Q_initial # Q_initial = 4.5
      
```

```{r, echo = FALSE }

num_options = 2
horizon = 100
group_size = 10
Q_initial = 4.5

      Q <- array(dim = c(num_options, horizon, group_size))
      Q[,1,] <- Q_initial # Q_initial = 4.5
      Q[,1:10,1] 
```

<br> **Global action probability**

Setup initial p

```{r, eval = FALSE}

netChoiceProb <- array(dim = c(num_options, horizon, group_size))
netChoiceProb[,1,] <- 1/num_options

      
```

```{r, echo = FALSE }

num_options = 2
horizon = 100
group_size = 10
Q_initial = 4.5

netChoiceProb <- array(dim = c(num_options, horizon, group_size))
netChoiceProb[,1,] <- 1/num_options
netChoiceProb[,1:10,1]
      
```

<br> **Action, reward and action probability**

```{r, eval = FALSE}


choices <- matrix(nrow=group_size, ncol=horizon)
payoffs <- matrix(nrow=group_size, ncol=horizon)
bestChoiceProb <- matrix(nrow=group_size, ncol=horizon)

```

<br> **Social frequency**

```{r, eval = FALSE}


# social frequency matrix to track available social information in the environment
      socialFrequency = matrix(nrow=num_options, ncol=horizon)
   
```

End of setup

<br>

### Step 1, t = 1

**Action**

Each agent chooses one option based on their choice probability matrix, and update their choices in the action matrix.

```{r, eval = FALSE}

t <- 1
   # A session starts for t in 1 : horizon
      for (t in 1:horizon) {
        # each agent chooses one option 
        choices[,t] <- apply(netChoiceProb[,t,]
                             , MARGIN = 2
                             , function(netChoiceProbInstance){ 
                               sample(1:num_options
                                      , 1
                                      , prob=netChoiceProbInstance
                                      , replace=FALSE
                               ) 
                             }
        )

```

Updated action matrix looks like:

```{r, echo = FALSE }

num_options = 2
horizon = 100
group_size = 10
Q_initial = 4.5

choices <- matrix(nrow=group_size, ncol=horizon)
choices[,1] <- c(2, 1, 1, 2, 2 ,2 ,2, 1, 1, 1)
choices[,1:10]      
```

**Reward**

Record the reward of the action of this step

```{r, eval = FALSE}

payoffs[,t] <- mapply(rnorm, 1, mean_list[choices[,t]], sd_list[choices[,t]])

```

Updated reward matrix looks like:

```{r, echo = FALSE }

num_options = 2
horizon = 100
group_size = 10
Q_initial = 4.5

payoffs <- matrix(nrow=group_size, ncol=horizon)
payoffs[,1] <- c(3.893331, 4.845328, 4.433069, 4.300975, 4.193934, 6.208168, 5.733692, 5.199272, 5.408271, 4.571809)

payoffs[,1:10]      

```

**Value**

Value is then updated by 

$$
Q_{a,t+1 }\leftarrow Q_{a,t } + \alpha\pi_{a,t}
$$



```{r, eval = FALSE}

           Q <- rescorla_wagner_updating(t
                                            , numoptions
                                            , group_size
                                            , horizon
                                            , this_alpha
                                            , choices[,t]
                                            , payoffs[,t]
                                            , Q
              )
              

```

Updated value matrix looks like:

```{r, echo = FALSE }

num_options = 2
horizon = 100
group_size = 10
Q_initial = 4.5

      Q <- array(dim = c(num_options, horizon, group_size))
      Q[,1,] <- Q_initial # Q_initial = 4.5
      Q[,2,1] <- c( 4.500000, 4.083182) # Q_initial = 4.5
      Q[,1:10,1]     
  
```

**Action probability**

Value is updated by $P_{a,t+1 } = (1-\sigma)\frac{exp(\beta Q_{a,t})}{\sum exp(\beta Q_{k,t})}+\sigma\frac{N_{a,t}^\theta}{\sum N_{k,t}^\theta}$

Asocial p:

```{r, eval = FALSE}
     
              # Calculate each "exp( beta*Q_i )"
              Q_exp = ( Q[,t+1,] * rep(this_beta, each = num_options) ) %>% exp() 
              # Then, calculating the softmax probability
              # exp( beta*Q_k )/(exp( beta*Q_1 ) + exp( beta*Q_2 )) for each slot
              softmaxMatrix = Q_exp %>% 
                apply(MARGIN=1
                      , divideVector
                      , denominator = apply(Q_exp, MARGIN=2, sum)
                ) %>% 
                t()
              
```

Social p:

1st, record social frequency based on the adjacency matrix of this time step

```{r, eval = FALSE}

 # read all matrices by col
              # Take the choice by col of this step and convert to a n*n matrix
              choiceMatrixThisT <- matrix(rep(choices[,t], times = group_size),nrow = group_size, ncol = group_size)
              # Choice of this step*graph of this step, get social frequency of each node by col
              SocialChoiceMat<- hadamard.prod(g, choiceMatrixThisT)
              
              # Count the social frequency for each node this step, matrix is this step, 
              # 0 is the lable of no SocialInfo, number bigger than 0 are choice labels
              FrequencyThisT <- apply(SocialChoiceMat, 2, function(x) table(factor(x,levels=1:num_options)))
              
              # record this step to both thisT and NextT, use accordingly
              # In this sim, nodes sample social frequency from t-1 step so use socialFrequencyNextT in later updating
              socialFrequencyThisT[,,t] <- socialFrequencyThisT[,,t] + FrequencyThisT # The real frequncy 
              socialFrequencyNextT[,,t+1] <- socialFrequencyNextT[,,t+1] + FrequencyThisT # The perceived frequency in this sim

```

2nd, calculate the social p

```{r, eval = FALSE}

       # ------ Calculating the social influence --------
              
              freqDepenMatrix <- 
                SN_frequencyDependent(socialFrequencyNextT[,,t+1]
                                      , choices[,t]
                                      , this_theta
                                      , num_options
                )

```

Asocial + social

```{r, eval = FALSE}

              netMatrix <- 
                apply(softmaxMatrix, 1, multiplying, A=(1-this_sigma)) %>%
                t() + 
                apply(freqDepenMatrix, 1, multiplying, A=this_sigma) %>% 
                t()

              netChoiceProbAperm = aperm(netChoiceProb, c(1,3,2))
              dim(netChoiceProbAperm) = c(num_options*group_size, horizon)
              dim(netMatrix) = c(num_options*group_size, 1)
              netChoiceProbAperm[,t+1] = netMatrix
              dim(netChoiceProbAperm) = c(num_options, group_size, horizon)
              netChoiceProb = aperm(netChoiceProbAperm, c(1,3,2))


```

Updated global action probability matrix looks like:

```{r, echo = FALSE }

num_options = 2
horizon = 100
group_size = 10
Q_initial = 4.5

netChoiceProb <- array(dim = c(num_options, horizon, group_size))
netChoiceProb[,1,] <- 1/num_options
netChoiceProb[,2,1] <- c( 0.9891614 , 0.0108386)
netChoiceProb[,1:10,1]

```

End of first step.

### Step 2, t = 2

Loop t

End of 100 step.

**Record this step** Summarize this step:

```{r, eval = FALSE}
best_mean <- which.max(mean_list)
          
          for(i in 1:group_size) {
            bestMeanProb[i,] <- netChoiceProb[best_mean,,i]
          }
          thisGroupPerformance = apply(bestMeanProb, MARGIN = 2, mean)
          thisGroupPayoff = apply(payoffs, MARGIN = 2, mean)  
          # Second, let's record this in the data set you defined 
          # We first find a position where new data should be recorded
          positionOfThisRun = which(social_learning_model_data$beta == beta & 
                                      social_learning_model_data$alpha == alpha &
                                      social_learning_model_data$sigma == sigma & 
                                      social_learning_model_data$theta == theta & 
                                      social_learning_model_data$r == r) 
```

Record it:

```{r, eval = FALSE}

          social_learning_model_data$bestMeanProb[positionOfThisRun] = thisGroupPerformance
          social_learning_model_data$averageReward[positionOfThisRun] = thisGroupPayoff
```

**Final results**

```{r, eval = FALSE}
head(social_learning_model_data)
```

```{r, echo = FALSE}
social_learning_model_data[1:6,7] <- c(0.5000000, 0.5784940, 0.6244481, 0.6990323, 0.7338047, 0.7176836)
social_learning_model_data[1:6,8] <- c(4.366376, 4.238600, 4.126742, 4.777364, 4.410486, 4.919184)
social_learning_model_data[1:6,] 
```

          

<br>

<br>


### Plot

```{r, include=FALSE}


library(ggplot2)
library(tidyverse)
library(dplyr)

fileP = getSourceEditorContext()$path
getP <- gsub("ESLR202402.Rmd", "data.csv", fileP)


dataset <- read.csv(getP) # Replace with the actual path or URL

ui <- fluidPage(
  
  
   sidebarLayout(

    # Sidebar with a slider input
    sidebarPanel(

    sliderInput("alpha_adjust", label = "alpha:",
              min = 0.2, max = 0.8, value = 0.2, step = 0.1),

    sliderInput("beta_adjust", label = "beta:",
              min = 2, max = 8, value = 2, step = 1)
    # 
    # sliderInput("sigma_adjust", label = "sigma:",
    #           min = 0, max = 1, value = 1, step = 0.1),
    # 
    # sliderInput("theta_adjust", label = "theta:",
    #           min = 0.01, max = 1, value = 0.5, step = 0.01)
),
   mainPanel(
  plotOutput("plot")

   )
))



# Define server logic
server <- function(input, output, session) {

  # Reactive dataset filtering
  filteredData <- reactive({
    req(input$alpha_adjust, input$beta_adjust)  # Ensure inputs are provided
    dataset %>%
      filter(beta == input$beta_adjust &  alpha == input$alpha_adjust)
  })
  
 
  # Create the plot
output$plot <- renderPlot({
  
  
    req(filteredData())
  
    ggplot(filteredData()) +
    geom_hline(yintercept = 0.5, linetype = 'dashed') +
    geom_line(aes(t, riskyChoiceProb_globalMean, colour = as.factor(sigma))) +
    scale_colour_viridis_d() +
    ylim(c(0,1))+
    labs(x = 'Steps'
         , y = '% Risky action'
         , colour = 'Inverse\nTemperature'
         , title = 'Social learning model') +
    facet_grid(. ~ theta) +
    theme_test() 


  })

}


shinyApp(ui = ui, server = server)



```

```{r, echo=FALSE}

library(shiny)
library(ggplot2)
library(tidyverse)
library(dplyr)

dataset <- read.csv("/Users/collectiveintelligence/Desktop/ESLR2024/RL/data.csv") # Replace with the actual path or URL
  



ui <- fluidPage(
  
  
   sidebarLayout(

    # Sidebar with a slider input
    sidebarPanel(

    # sliderInput("alpha_adjust", label = "alpha:",
    #           min = 0.2, max = 0.8, value = 0.2, step = 0.1),
    # 
    # sliderInput("beta_adjust", label = "beta:",
    #           min = 2, max = 8, value = 2, step = 1)

    sliderInput("sigma_adjust", label = "sigma:",
               min = 0, max = 1, value = 0.6, step = 0.2),

    sliderInput("theta_adjust", label = "theta:",
              min = 1, max = 7, value = 3, step = 2)
),
   mainPanel(
  plotOutput("plot")

   )
))



# Define server logic
server <- function(input, output, session) {

  # Reactive dataset filtering
  filteredData <- reactive({
    req(input$sigma_adjust, input$theta_adjust)  # Ensure inputs are provided
    dataset %>%
      filter(sigma == input$sigma_adjust &  theta == input$theta_adjust)
  })
  
 
  # Create the plot
output$plot <- renderPlot({
  
  
    req(filteredData())
  
    ggplot(filteredData()) +
    geom_hline(yintercept = 0.5, linetype = 'dashed') +
    geom_line(aes(t, riskyChoiceProb_globalMean, colour = as.factor(beta))) +
    scale_colour_viridis_d() +
    ylim(c(0,1))+
    labs(x = 'Steps'
         , y = '% Optimal action'
         , colour = 'Inverse\nTemperature'
         , title = 'Social learning model') +
    facet_grid(. ~ alpha) +
    theme_test() 


  })

}


shinyApp(ui = ui, server = server)



```

## References

Toyokawa, W., Whalen, A., & Laland, K. N. (2019). Social learning strategies regulate the wisdom and madness of interactive crowds. Nature human behaviour, 3(2), 183-193.

Wilson, R. C., & Collins, A. G. (2019). Ten simple rules for the computational modeling of behavioral data. Elife, 8, e49547.

Sutton, R. S. (2018). Reinforcement learning: An introduction. A Bradford Book.

Newman, M. (2018). Networks. Oxford university press.

Barabási, A. L. (2013). Network science. Philosophical Transactions of the Royal Society A: Mathematical, Physical and Engineering Sciences, 371(1987), 20120375.
